---
title: "Statistical Inference Project 2"
author: "JD"
date: "Saturday, May 23, 2015"
graphics: yes
output: pdf_document 
---
#### Overview
The objective 
To analyze the ToothGrowth data in the R datasets package ToothGrowth. This has observations for 60 samples (Animals). These animals were divided into 6 groups of 10 and consistently fed a diet with one of 6 Vitamin C supplement regimes for a period of 42 days. The Vitamin C was administered either in the form of Orange Juice (OJ) or chemically pure Vitamin C in aqueous solution (VC). Each animal received the same daily dosage of Vitamin C (either 0.5, 1.0 or 2.0 milligrams) consistently.  

```{r global_options , echo=FALSE, fig.align='left'}

```

#### 1. Load the ToothGrowth data 
```{r}
library(ggplot2)
library(datasets)
data(ToothGrowth)

## This function returns the title for facet grids (Ref http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/)
mf_labeller <- function(var, value){
  value <- as.character(value)
  if (var=="supp") { 
    value[value=="OJ"] <- "Orange Juice"
    value[value=="VC"]   <- "Vit C Aqeuous Soln"
  }
  return(value)
}
```
_**Displaying Tooth Growth Datasets basics**_

#### 2. Summary and Structure

**A. Summary**
```{r}
summary(ToothGrowth)
```

**B. Basic Structure**
```{r}
str(ToothGrowth)
```

#### 3. Displaying Tooth Growth Plot and basic analysis

```{r, fig.align='left'}

ggplot(data=ToothGrowth, aes(dose, len)) +geom_bar(stat="identity") + facet_grid(.~supp, drop=FALSE,labeller=mf_labeller) + xlab("DOSE") + ylab("Tooth Length")+
  ggtitle("Plot of Tooth Length vs Dosage of the two Supplements ")

```



#### 4. Initial Inference

_The initial analysis shows a direct relationship between the dosage and tooth growths. At lower dosages (0.5 and 1.0) OJ seemed to provide better tooth growth then with Vit C supplements. OJ provides a 50 micron growth over the Vit C supplement._ 
_At a dosage of 2.0 units though both supplements provide a comparable level of tooth growth._


#### 5. Detailed Analysis

**Here we attain p values to evaluate null hypothesis.**

**A. Load Datasets for dosage levels**
```{r}
## Load data sets for the three dosage levels
dosage01 <- subset(ToothGrowth, dose==0.5)
dosage02 <- subset(ToothGrowth, dose==1.0)
dosage03 <- subset(ToothGrowth, dose==2.0)

```

**B. Perform T Test on the three sets of data. P Values printed below**
```{r}
t<-t.test(len~supp, paired=FALSE, data=dosage01)
t$p.value
t<-t.test(len~supp, paired=FALSE, data=dosage02)
t$p.value
t<-t.test(len~supp, paired=FALSE, data=dosage03)
t$p.value
```

#### 6. Analysis

_Dosage of 0.5 and 1.0 have p values of 0.0063 and 0.0010 respectively. Dosage of 2.0 has p value of 0.9638._
_Orange Juice and Vit C has greater impact on tooth growth at lower levels. At higher dosages the differences caused by the two supplements have no differences to tooth growth._



#### 7. References

[Refer This Link for details on P Value ( http://www.dummies.com/how-to/content/what-a-pvalue-tells-you-about-statistical-data.html)


